The Proofs

### 2024-02-13 proofs/main.tex
wip

Let us discuss the selected rules of the declarative system:
\begin{itemize}
  \item \ruleref{\ottdruleDTVarLabel}
    says that the type of a variable is inferred from the context.
  \item \ruleref{\ottdruleDTThunkLabel} says
    that the type of a thunk is inferred by shifting up the type of the 
    contained computation. Symmetrically, \ruleref{\ottdruleDTReturnLabel}
    infers the type of a return by shifting down the type of the
    contained value.
  \item \ruleref{\ottdruleDTPAnnotLabel} and \ruleref{\ottdruleDTNAnnotLabel} are symmetric.
    They allow the inferred type to be refined by annotating it with a supertype.
  \item \ruleref{\ottdruleDTNEquivLabel} and \ruleref{\ottdruleDTPEquivLabel}
    mean that the declarative system allows us to infer any type from the equivalence class.
  \item \ruleref{\ottdruleDTUnpackLabel} is standard for existential types,
    and its first premise infers the existential type of the value being unpacked.
    It is important however that the inferred existential type is normalized. 
    This is because there might be multiple equivalent existential types 
    with a different order or even number of quantified variables, 
    and to bind them, the algorithm needs to fix the canonical one.
  \item \ruleref{\ottdruleDTAppLetAnnLabel} allows us to accommodate the applications
    with annotated let-bindings. The first premise infers the type of the head of the application,
    which must be a thunked computation. Then if after applying it 
    to the arguments, the resulting type can be equated to the annotated one,
    we infer the body of the let-binding in the context extended with the bound variable.
  \item \ruleref{\ottdruleDTAppLetLabel} is similar to \ruleref{\ottdruleDTAppLetAnnLabel},
    but it is used when the type of the application is unique, and thus, the annotation
    is redundant. Here $[[Γ ; Φ ⊢ iM ● args ⇒> ↑iQ uniq]]$ means that
    if also $[[Γ ; Φ ⊢ iM ● args ⇒> ↑iQ']]$ then $[[Γ ⊢ iQ ≈ iQ']]$
\end{itemize}

@@

Let us discuss the inference rules of the algorithm:
\begin{itemize}
  \item \ruleref{\ottdruleATVarLabel} 
    infers the type of a variable by looking it up in the context
    and normalizing the result.
  \item \ruleref{\ottdruleATThunkLabel} and \ruleref{\ottdruleATReturnLabel}
    are similar to the declarative rules: they make a recursive call
    to type the body of the thunk or the return expression and
    put the shift on top of the result.
  \item \ruleref{\ottdruleATPAnnotLabel} and \ruleref{\ottdruleATNAnnotLabel}
    are symmetric. They make a recursive call to infer the type of the annotated
    expression, check that the inferred type is a subtype of the annotation,
    and return the normalized annotation.
  \item \ruleref{\ottdruleATtLamLabel} infers the type of a lambda-abstraction.
    It makes a recursive call to infer the type of the body in the extended context,
    and returns the corresponding arrow type. Notice that the algorithm also
    normalizes the result, which is because the annotation type $[[iP]]$
    is allowed to be non-normalized.
  \item \ruleref{\ottdruleATTLamLabel} infers the type of a big lambda.
    Similarly to the previous case, it makes a recursive call to infer the type
    of the body in the extended \emph{type} context. 
    After that, it returns the corresponding universal type. 
    It is also required to normalize the result, because, 
    for instance, $[[α⁺]]$ might not occur in the body of the lambda,
    in which case the $[[∀]]$ must be removed. 
  \item \ruleref{\ottdruleATVarLetLabel} is defined in a standard way:
    it makes a recursive call to infer the type of the bound value,
    and then returns the type of the body in the extended context.
  \item \ruleref{\ottdruleATAppLetAnnLabel}
    is interpreted as follows.
    First, it infers the type of the head of the application,
    ensuring that it is a thunked computation $[[↓iM]]$;
    after that, it makes a recursive call
    to the application inference procedure,
    which returns the algorithmic type, whose
    instantiation to a declarative type must be associated with the bound variable 
    $[[x]]$; then premise $[[Γ; Θ ⊨ ↑uQ ≤ ↑iP ⫤ SC2]]$
    together with $[[Θ ⊢ SC1 & SC2 = SC]]$
    check whether the instantiation to the annotated type $[[iP]]$ is possible,
    and if it is, the algorithm infers the type of the body in the extended context,
    and returns it as the result. 
  \item \ruleref{\ottdruleATAppLetAnnLabel}
    works similarly to \ruleref{\ottdruleATAppLetAnnLabel},
    However, since there is no annotation, 
    instead of checking the instantiation to it, 
    the algorithm checks that the inferred type
    $[[↑uQ]]$ is unique.
    It is the case if all the algorithmic variables of 
    $[[↑uQ]]$ are sufficiently restricted by $[[SC]]$,
    which is checked by the combination of
    $[[uv uQ = dom(SC)]]$ and $[[SC singular with uσ]]$.
    Together, these two premises guarantee that the only 
    possible instantiation of $[[↑uQ]]$ is $[[ [uσ]uQ ]]$.
  \item \ruleref{\ottdruleATUnpackLabel}
    works in the expected way. First, it infers the 
    existential type $[[∃nas.iP]]$ of the value being unpacked,
    and since the type is guaranteed to be normalized, binds 
    the quantified variables with $[[nas]]$.
    Then it infers the type of the body in the appropriately extended context,
    and checks that the inferred type does not depend on $[[nas]]$
    by checking well-formedness $[[Γ ⊢ iN]]$.
\end{itemize}

@@

@@

\subsection{Singularity}
\input{\genDir/tex/singularity-statements.tex}

@@

\subsection{Singularity}
\input{\genDir/tex/singularity-lemmas.tex}

>>>>

Let us discuss the selected rules of the declarative system:
\begin{itemize}
  \item \ruleref{\ottdruleDTVarLabel}
    says that the type of a variable is inferred from the context.
  \item \ruleref{\ottdruleDTThunkLabel} says
    that the type of a thunk is inferred by shifting up the type of the 
    contained computation. Symmetrically, \ruleref{\ottdruleDTReturnLabel}
    infers the type of a return by shifting down the type of the
    contained value.
  \item \ruleref{\ottdruleDTPAnnotLabel} and \ruleref{\ottdruleDTNAnnotLabel} are symmetric.
    They allow the inferred type to be refined by annotating it with a supertype.
  \item \ruleref{\ottdruleDTNEquivLabel} and \ruleref{\ottdruleDTPEquivLabel}
    mean that the declarative system allows us to infer any type from the equivalence class.
  \item \ruleref{\ottdruleDTUnpackLabel} is standard for existential types,
    and its first premise infers the existential type of the value being unpacked.
    It is important however that the inferred existential type is normalized. 
    This is because there might be multiple equivalent existential types 
    with a different order or even number of quantified variables, 
    and to bind them, the algorithm needs to fix the canonical one.
  \item \ruleref{\ottdruleDTAppLetAnnLabel} allows us to accommodate the applications
    with annotated let-bindings. The first premise infers the type of the head of the application,
    which must be a thunked computation. Then if after applying it 
    to the arguments, the resulting type can be equated to the annotated one,
    we infer the body of the let-binding in the context extended with the bound variable.
  \item \ruleref{\ottdruleDTAppLetLabel} is similar to \ruleref{\ottdruleDTAppLetAnnLabel},
    bus is used for unannotated let-bindings. In this case, 
    we require the type application to infer the `canonical' principal type.
    $[[Γ ; Φ ⊢ iM ● args ⇒> ↑iQ principal]]$ means that
    any other type $[[iQ']]$ inferrable for the application (\ie $[[Γ ; Φ ⊢ iM ● args ⇒> ↑iQ']]$)
    is greater than the principal type $[[iQ]]$, \ie $[[Γ ⊢ iQ' ≥ iQ]]$.
\end{itemize}

@@

Let us discuss the inference rules of the algorithm:
\begin{itemize}
  \item \ruleref{\ottdruleATVarLabel} 
    infers the type of a variable by looking it up in the context
    and normalizing the result.
  \item \ruleref{\ottdruleATThunkLabel} and \ruleref{\ottdruleATReturnLabel}
    are similar to the declarative rules: they make a recursive call
    to type the body of the thunk or the return expression and
    put the shift on top of the result.
  \item \ruleref{\ottdruleATPAnnotLabel} and \ruleref{\ottdruleATNAnnotLabel}
    are symmetric. They make a recursive call to infer the type of the annotated
    expression, check that the inferred type is a subtype of the annotation,
    and return the normalized annotation.
  \item \ruleref{\ottdruleATtLamLabel} infers the type of a lambda-abstraction.
    It makes a recursive call to infer the type of the body in the extended context,
    and returns the corresponding arrow type. Notice that the algorithm also
    normalizes the result, which is because the annotation type $[[iP]]$
    is allowed to be non-normalized.
  \item \ruleref{\ottdruleATTLamLabel} infers the type of a big lambda.
    Similarly to the previous case, it makes a recursive call to infer the type
    of the body in the extended \emph{type} context. 
    After that, it returns the corresponding universal type. 
    It is also required to normalize the result, because, 
    for instance, $[[α⁺]]$ might not occur in the body of the lambda,
    in which case the $[[∀]]$ must be removed. 
  \item \ruleref{\ottdruleATVarLetLabel} is defined in a standard way:
    it makes a recursive call to infer the type of the bound value,
    and then returns the type of the body in the extended context.
  \item \ruleref{\ottdruleATAppLetAnnLabel}
    is interpreted as follows.
    First, it infers the type of the head of the application,
    ensuring that it is a thunked computation $[[↓iM]]$;
    after that, it makes a recursive call
    to the application inference procedure,
    which returns the algorithmic type, whose
    instantiation to a declarative type must be associated with the bound variable 
    $[[x]]$; then premise $[[Γ; Θ ⊨ ↑uQ ≤ ↑iP ⫤ SC2]]$
    together with $[[Θ ⊢ SC1 & SC2 = SC]]$
    check whether the instantiation to the annotated type $[[iP]]$ is possible,
    and if it is, the algorithm infers the type of the body in the extended context,
    and returns it as the result. 
  \item \ruleref{\ottdruleATAppLetLabel}
    works similarly to \ruleref{\ottdruleATAppLetAnnLabel},
    However, since there is no annotation to assign the result to,
    the algorithm must infer the `canonical' principal type.
    To do that algorithmically, we ensure that 
    the inferred algorithmic type $[[uQ]]$ is instantiated
    to the minimal possible type $[[ [uσ]uQ ]]$.
    The premise $[[Γ ⊢ uQ SC minby uσ]]$
    provides the minimal instantiation of $[[uQ]]$ \wrt $[[SC]]$. It 
    guarantees that if we consider all possible substitutions satisfying the inferred 
    constraints $[[SC]]$, then substitution $[[uσ]]$ will instantiate
    $[[uQ]]$ to the minimal possible type $[[ [uσ]uQ ]]$.
    This will be the principal type that we assign to the result of the application 
    (bound to the variable $[[x]]$)
    and then we infer the type of the body in the context extended with 
    the bound variable $[[x : [uσ]uQ]]$.
  \item \ruleref{\ottdruleATUnpackLabel}
    first, infers the existential type $[[∃nas.iP]]$ of the value being unpacked,
    and since the type is guaranteed to be normalized, binds 
    the quantified variables with $[[nas]]$.
    Then it infers the type of the body in the appropriately extended context
    and checks that the inferred type does not depend on $[[nas]]$
    by checking well-formedness $[[Γ ⊢ iN]]$.
\end{itemize}

@@

\subsection{Minimal Instantiation}
The minimal instantiation algorithm is used
to infer the type of the bound variable in the un-annotated 
applicative let-binders, as long as there exists a principal
(minimal) type.  
Given a positive algorithmic type $[[uP]]$ 
and a set of constraints $[[SC]]$, it finds the substitution $[[uσ]]$
respecting $[[SC]]$ (i.e., $[[Θ ⊢ uσ : SC]]$) such that
it instantiates $[[uP]]$ to the minimal type, in other words 
for any other substitution $[[uσ']]$ respecting $[[SC]]$,
we have $[[Γ ⊢ [uσ']uP ≥ [uσ]uP]]$.

The minimal instantiation algorithm is defined as follows:
\begin{algorithm}[Minimal Instantiation]
  \hfill \\
  \ottdefnSINGminInstLabeled{}
\end{algorithm}

@@

\subsection{Singularity and Minimal Instantiation}
\input{\genDir/tex/singularity-statements.tex}

@@

\subsection{Singularity and Minimal Instantiation}
\input{\genDir/tex/singularity-lemmas.tex}

### 2024-02-13 proofs/tex/decl-typing-lemmas.tex
wip

              And after that, we apply the same rule to infer the required judgement.

@@

        \item \ruleref{\ottdruleDTtLamLabel}
            Then we are proving that 
            $[[Γ ; Φ1 ⊢ λx:iP.c : iP → iN]]$ implies $[[Γ ; Φ2 ⊢ λx:iP.c : iP → iN]]$.
            Analogously to the previous cases, 
            we apply the induction hypothesis to the
            equivalent contexts $[[Γ ⊢ Φ1, x:iP ≈ Φ2, x:iP]]$
            and the premise $[[Γ ; Φ1, x:iP ⊢ c : iN]]$
            to obtain $[[Γ ; Φ2, x:iP ⊢ c : iN]]$.
            Notice that $[[c]]$ is a subterm of $[[λx:iP.c]]$,
            i.e., the metric of the premise tree is less than the metric of the conclusion, 
            and the induction hypothesis is applicable.
            Then we infer $[[Γ ; Φ2 ⊢ λx:iP.c : iP → iN]]$ by \ruleref{\ottdruleDTtLamLabel}.

        \item \ruleref{\ottdruleDTVarLetLabel}
            Then we are proving that 
            $[[Γ ; Φ1 ⊢ let x = v; c : iN]]$ implies $[[Γ ; Φ2 ⊢ let x = v; c : iN]]$.
            First, we apply the induction hypothesis to 
            $[[Γ; Φ1 ⊢ v : iP]]$ to obtain $[[Γ; Φ2 ⊢ v : iP]]$ 
            of the same pure size.

@@

        \item \ruleref{\ottdruleDTAppLetLabel}
            Then we are proving that 
            $[[Γ ; Φ1 ⊢ let x = v(args); c : iN]]$ implies 
            $[[Γ ; Φ2 ⊢ let x = v(args); c : iN]]$.

@@

            It is left to show the uniqueness of $[[Γ ; Φ2 ⊢ iM ● args ⇒> ↑iQ]]$.
            Let us assume that this judgement holds for other $[[iQ']]$, 
            i.e.  there exists a tree $T_0$ inferring 
            $[[Γ ; Φ2 ⊢ iM ● args ⇒> ↑iQ']]$.
            Then notice that the induction hypothesis is applicable to
            $T_0$: the first component of the first component of $\metric{T_0}$
            is $S = \sum_{[[v]] \in [[args]]} \size{[[v]]}$, and it is less than
            the corresponding component of $\metric{T_1}$, which is
            $\size{[[let x = v(args); c]]} = 1 + \size{[[v]]} + \size{[[c]]} + S$.
            This way, $[[Γ ; Φ1 ⊢ iM ● args ⇒> ↑iQ']]$ holds by the induction hypothesis,
            but since $[[Γ ; Φ1 ⊢ iM ● args ⇒> ↑iQ uniq]]$, we have $[[Γ ⊢ iQ' ≈ iQ]]$.

            Then we infer $[[Γ ; Φ2 ⊢ let x = v(args); c : iN]]$ by \ruleref{\ottdruleDTAppLetLabel}.

        \item \ruleref{\ottdruleDTAppLetAnnLabel}
            Then we are proving that
            $[[Γ ; Φ1 ⊢ let x:iP = v(args); c : iN]]$ implies
            $[[Γ ; Φ2 ⊢ let x:iP = v(args); c : iN]]$.

>>>>

              And after that, we apply the same rule to infer the required judgment.

@@

        \item \ruleref{\ottdruleDTtLamLabel}
            Then we prove that 
            $[[Γ ; Φ1 ⊢ λx:iP.c : iP → iN]]$ implies $[[Γ ; Φ2 ⊢ λx:iP.c : iP → iN]]$.
            Analogously to the previous cases, 
            we apply the induction hypothesis to the
            equivalent contexts $[[Γ ⊢ Φ1, x:iP ≈ Φ2, x:iP]]$
            and the premise $[[Γ ; Φ1, x:iP ⊢ c : iN]]$
            to obtain $[[Γ ; Φ2, x:iP ⊢ c : iN]]$.
            Notice that $[[c]]$ is a subterm of $[[λx:iP.c]]$,
            i.e., the metric of the premise tree is less than the metric of the conclusion, 
            and the induction hypothesis is applicable.
            Then we infer $[[Γ ; Φ2 ⊢ λx:iP.c : iP → iN]]$ by \ruleref{\ottdruleDTtLamLabel}.

        \item \ruleref{\ottdruleDTVarLetLabel}
            Then we prove that 
            $[[Γ ; Φ1 ⊢ let x = v; c : iN]]$ implies $[[Γ ; Φ2 ⊢ let x = v; c : iN]]$.
            First, we apply the induction hypothesis to 
            $[[Γ; Φ1 ⊢ v : iP]]$ to obtain $[[Γ; Φ2 ⊢ v : iP]]$ 
            of the same pure size.

@@

        \item \ruleref{\ottdruleDTAppLetLabel}
            Then we prove that 
            $[[Γ ; Φ1 ⊢ let x = v(args); c : iN]]$ implies 
            $[[Γ ; Φ2 ⊢ let x = v(args); c : iN]]$.

@@

            It is left to show the principality of $[[Γ ; Φ2 ⊢ iM ● args ⇒> ↑iQ]]$.
            Let us assume that this judgment holds for other $[[iQ']]$, 
            i.e.  there exists a tree $T_0$ inferring 
            $[[Γ ; Φ2 ⊢ iM ● args ⇒> ↑iQ']]$.
            Then notice that the induction hypothesis applies to
            $T_0$: the first component of the first component of $\metric{T_0}$
            is $S = \sum_{[[v]] \in [[args]]} \size{[[v]]}$, and it is less than
            the corresponding component of $\metric{T_1}$, which is
            $\size{[[let x = v(args); c]]} = 1 + \size{[[v]]} + \size{[[c]]} + S$.
            This way, $[[Γ ; Φ1 ⊢ iM ● args ⇒> ↑iQ']]$ holds by the induction hypothesis,
            but since $[[Γ ; Φ1 ⊢ iM ● args ⇒> ↑iQ]]$ is principal, 
            we have $[[Γ ⊢ iQ' ≥ iQ]]$. 
            This way, $[[Γ ; Φ2 ⊢ iM ● args ⇒> ↑iQ principal]]$.
            Then we infer $[[Γ ; Φ2 ⊢ let x = v(args); c : iN]]$ by \ruleref{\ottdruleDTAppLetLabel}.

        \item \ruleref{\ottdruleDTAppLetAnnLabel}
            Then we prove that
            $[[Γ ; Φ1 ⊢ let x:iP = v(args); c : iN]]$ implies
            $[[Γ ; Φ2 ⊢ let x:iP = v(args); c : iN]]$.

### 2024-02-13 proofs/tex/singularity-lemmas.tex
wip

>>>>

\lemMinInstSoundness*
\begin{proof}
    We prove it by induction on the inference of $[[Γ ⊢ uP SC minby uσ ]]$.
    Let us consider the last rule used in the inference.
    \begin{caseof}
        \item \ruleref{\ottdruleSINGPUvarLabel}, which means 
            that the inferred judgment is $[[Γ ⊢  α̂⁺ SC minby (nf(iP) / α̂⁺)]]$, 
            and by inversion, $[[(α̂⁺ :≥ iP) ∊ SC]]$. Let us show the required properties:
            \begin{itemize}
                \item $[[Θ ⊢ (nf(iP) / α̂⁺) : uv α̂⁺]]$ holds trivially;
                \item $[[Θ ⊢  (nf(iP) / α̂⁺) : SC]]$ holds since
                    $[[Γ ⊢ nf(iP) : (α̂⁺ :≥ iP)]]$, which is true 
                    since $[[Γ ⊢ nf(iP) ≥ iP]]$ by the soundness of normalization
                    (\cref{lemma:normalization-soundness});
                \item $[[(nf(iP) / α̂⁺)]]$ is normalized trivially;
                \item let us take an arbitrary $[[Θ ⊢ uσ' : {α̂⁺} ]]$
                    respecting $[[SC]]$. Since $[[uσ']]$ respects $[[SC]]$,
                    $[[Γ ⊢ [uσ']α̂⁺ ≥ iP]]$ holds, and then
                    $[[Γ ⊢ [uσ']α̂⁺ ≥ nf(iP)]]$ holds by the soundness of normalization
                    and transitivity of subtyping. Finally,
                    $[[Γ ⊢ [uσ']α̂⁺ ≥ nf(iP)]]$ can be rewritten as
                    $[[Γ ⊢ [uσ']α̂⁺ ≥ [(nf(iP) / α̂⁺)]α̂⁺]]$.
            \end{itemize}
        \item \ruleref{\ottdruleSINGExistsLabel}, which means that 
            the inferred judgment has form $[[Γ ⊢ ∃nas.uP SC minby uσ ]]$,
            and by inversion, $[[Γ, nas ⊢ uP SC minby uσ]]$.
            By applying the induction hypothesis to $[[Γ, nas ⊢ uP SC minby uσ]]$
            we have 
            \begin{itemize}
                \item $[[Θ ⊢ uσ : uv uP]]$, which also means $[[Θ ⊢ uσ : uv ∃nas.uP]]$,
                \item $[[Θ ⊢ uσ : SC]]$,
                \item $[[uσ]]$ is normalized, and
                \item for any other $[[Θ ⊢ uσ' : uv uP ]]$
                    respecting $[[SC]]$ (\ie $[[Θ ⊢ uσ' : SC]]$), we have
                    $[[Γ, nas ⊢ [uσ']uP ≥ [uσ]uP ]]$, which immediately implies 
                    $[[Γ ⊢ [uσ']∃nas.uP ≥ [uσ]∃nas.uP ]]$ (the left-hand side
                    existential variables are instantiated with the corresponding 
                    right-hand side existential variables).
            \end{itemize}
        \item \ruleref{\ottdruleSINGSingLabel}, which means that
            the inferred judgment has form $[[Γ ⊢ uP SC minby uσ]]$,
            and by inversion, $[[uv(uP) ⊆ dom(SC)]]$ and
            $[[SC|uv(uP) singular with uσ]]$. Let us apply the soundness of singularity
            (\cref{lemma:singularity-soundness}) to $[[SC|uv(uP) singular with uσ]]$ to
            obtain the following properties:
            \begin{itemize}
                \item $[[ Θ ⊢ uσ : uv(uP) ∩ dom(SC) ]]$, which also means 
                    $[[Θ ⊢ uσ : uv(uP)]]$,
                \item $[[Θ ⊢ uσ : SC|uv(uP)]]$,
                \item $[[uσ]]$ is normalized, and
                \item for any other $[[Θ ⊢ uσ' : uv(uP) ]]$
                    respecting $[[SC|uv(uP)]]$, we have
                    $[[Θ ⊢ uσ' ≈ uσ : uv(uP)]]$. The latter means that
                    $[[Γ ⊢ [uσ']uP ≈ [uσ]uP]]$, and in particular, 
                    $[[Γ ⊢ [uσ']uP ≥ [uσ]uP]]$.
            \end{itemize}
    \end{caseof}

\lemMinInstCompleteness*
\begin{proof}
    We prove it by induction on $[[uP]]$.
    \begin{caseof}
        \item \label{case:min-inst-completeness-uvar} 
            $[[uP]] = [[α̂⁺]]$. Suppose that $[[α̂⁺ ∉ dom(SC)]]$.
            Then the instantiation of $[[α̂⁺]]$ is
            not restricted, and thus, any type can instantiate it.
            However, among unrestricted instantiations, there is no minimum:
            any type $[[iP]]$ is \emph{not} a subtype of $[[↓↑iP]]$,
            which contradicts the assumption.
            This way, $[[α̂⁺ ∊ dom(SC)]]$.

            If the entry restricting $[[α̂⁺]]$ in $[[SC]]$ is a \emph{subtyping} entry
            ($[[(α̂⁺ :≥ iQ) ∊ SC]]$), then 
            we apply \ruleref{\ottdruleSINGPUvarLabel} to infer 
            $[[Γ ⊢  α̂⁺ SC minby (nf(iQ) / α̂⁺)]]$.
            It is left to show that $[[ nf(iQ) = nf([uσ]α̂⁺) ]]$.
            Since $[[Θ ⊢ uσ : SC]]$, and $[[(α̂⁺ :≥ iQ) ∊ SC]]$, 
            we know that $[[Γ ⊢ [uσ]α̂⁺ ≥ iQ]]$. 
            On the other hand, let us consider $[[Θ ⊢ uσ' : SC]]$,
            that copies $[[uσ]]$ on $[[dom(SC)]]$ except $[[α̂⁺]]$,
            where it is instantiated with $[[iQ]]$.
            Then $[[Γ ⊢ [uσ']α̂⁺ ≥ [uσ]α̂⁺ ]]$ means $[[Γ ⊢ iQ ≥ [uσ]α̂⁺ ]]$,
            this way, $[[Γ ⊢ iQ ≈ [uσ]α̂⁺ ]]$, which by \cref{lemma:subt-equiv-algorithmization}
            means $[[nf(iQ) = nf([uσ]α̂⁺)]]$.

            If the entry restricting $[[α̂⁺]]$ in $[[SC]]$ is an \emph{equivalence} entry
            ($[[(α̂⁺ :≈ iQ) ∊ SC]]$), then we wish to apply
            \ruleref{\ottdruleSINGSingLabel}.
            The first premise $[[uv(α̂⁺) ⊆ dom(SC)]]$  
            holds by assumption;
            to infer $[[SC|{α̂⁺} singular with uσ0]]$, 
            we apply the completeness of singularity (\cref{lemma:singularity-completeness}).
            It applies because all the substitutions satisfying $[[SC|{α̂⁺}]] = [[(α̂⁺ :≈ iQ)]]$
            are equivalent on $[[{α̂⁺}]]$ by transitivity of equivalence 
            (\cref{corollary:equivalence-transitivity}): 
            the satisfaction of this constraint means that the substitution 
            sends $[[α̂⁺]]$ to $[[iQ]]$ or an equivalent type.
            This way, $[[SC|{α̂⁺} singular with uσ0]]$ for some $[[uσ0]]$,
            which means $[[Γ ⊢ α̂⁺ SC minby uσ0]]$.
            To show that $[[uσ0 = nf(uσ)]]$ notice that 
            Since $[[uσ0]]$ is normalized and equivalent to $[[uσ]]$ on $[[{α̂⁺}]]$, 
            and only has $[[α̂⁺]]$ in its domain
            (by soundness of singularity, \cref{lemma:singularity-soundness}).
            This way, $[[Γ ⊢ α̂⁺ SC minby uσ]]$, as required.
        \item  \label{case:min-inst-completeness-down}
            $[[uP = ↓uN]]$. Then since
            $[[Γ ⊢ ↓[uσ']uN ≥ ↓[uσ]uN ]]$ means
            $[[Γ ⊢ ↓[uσ']uN ≈ ↓[uσ]uN ]]$ by inversion.
            Then by \cref{lemma:subst-equiv}, 
            $[[uσ]]$ is equivalent to any other substitution 
            $[[Θ ⊢ uσ' : uv uN ]]$ satisfying $[[SC|uv uN]]$,
            hence, the completeness of singularity (\cref{lemma:singularity-completeness})
            can be applied to conclude that
            \begin{itemize}
                \item $[[uv(uN) = dom(SC|uv uN)]]$, then $[[uv(uP) ⊆ dom(SC)]]$,
                \item $[[SC|uv uN singular with uσ0]]$ for some (normalized) $[[uσ0]]$.
            \end{itemize}
            It means $[[Γ ⊢ uP SC minby uσ0 ]]$, and then
            since $[[uσ0]]$ is normalized and equivalent to $[[uσ0]]$ on $[[uv(uN)]]$,
            and its domain is $[[uv(uN)]]$, $[[uσ0 = nf(uσ)]]$.
        % \item $[[uP = ∃nas.uP']]$ 
        %     Then we wish to apply the induction hypothesis to show 
        %     $[[Γ, nas ⊢ uP' SC minby nf(uσ)]]$, which would imply 
        %     $[[Γ ⊢ ∃nas.uP' SC minby nf(uσ)]]$ by \ruleref{\ottdruleMINExistsLabel}.
        %     Let us demonstrate that the premises of the induction hypothesis hold. 
        %     \begin{itemize}
        %         \item  $[[Γ, nas ⊢ Θ]]$ holds by weakening;
        %         \item  $[[Γ, nas; dom(Θ) ⊢ uP']]$ holds by inversion of $[[Γ; dom(Θ) ⊢ uP]]$;
        %         \item $[[uv uP' = uv uP]]$, and 
        %                 $[[Γ ⊢ [uσ']∃nas.uP' ≥ [uσ]∃nas.uP' ]]$ implies
        %                 $[[Γ, nas ⊢ [uσ']uP' ≥ [uσ]uP' ]]$ by \cref{lemma:subst-equiv}.
        %     \end{itemize}

        \item $[[uP = ∃nas.β⁺]]$ then 
            as there are no algorithmic variables in $[[uP]]$, 
            $[[nf([uσ]uP) = β⁺]]$,
            and thus, we wish to show that 
            $[[Γ ⊢ ∃nas.β⁺ SC minby ·]]$.
            To do so, we apply \ruleref{\ottdruleSINGExistsLabel},
            and it is left to show that $[[Γ, nas ⊢ β⁺ SC minby · ]]$,
            which holds vacuously by \ruleref{\ottdruleSINGSingLabel}.
        \item $[[uP = ∃nas.α̂⁺]]$ then 
            $[[Γ ⊢ [uσ']∃nas.α̂⁺ ≥ [uσ]∃nas.α̂⁺ ]]$ implies 
            $[[Γ ⊢ ∃nas.[uσ']α̂⁺ ≥ ∃nas.[uσ]α̂⁺ ]]$ implies 
            $[[Γ ⊢ [uσ']α̂⁺ ≥ [uσ]α̂⁺ ]]$.
            It means that $[[uσ]]$ instantiates $[[α̂⁺]]$ to the minimal 
            type among all the instantiations of $[[α̂⁺]]$ respecting $[[SC]]$.
            In other words, we can apply the reasoning from 
            \cref{case:min-inst-completeness-uvar}
            to conclude that $[[Γ ⊢ α̂⁺ SC minby nf(uσ) ]]$.
            And then \ruleref{\ottdruleSINGExistsLabel}
            gives us $[[Γ ⊢ ∃nas.α̂⁺ SC minby nf(uσ) ]]$.
        \item $[[uP = ∃nas.↓uN]]$ then 
            $[[Γ ⊢ [uσ']∃nas.↓uN ≥ ∃nas.↓[uσ]uN ]]$ implies 
            $[[Γ ⊢ ∃nas.↓[uσ']uN ≥ ∃nas.↓[uσ]uN ]]$ implies
            $[[Γ, nas ⊢ ↓[σ0][uσ']uN ≥ ↓[uσ]uN ]]$ for some $[[σ0]]$ implies
            $[[Γ, nas ⊢ [σ0][uσ']uN ≈ [uσ]uN ]]$.
            By \cref{lemma:subst-equiv}, it means in particular
            that $[[uσ']]$ and $[[uσ]]$ are equivalent on 
            $[[uv uN]]$. 
            This way, we can apply the completeness of singularity 
            (\cref{lemma:singularity-completeness}), and continue as in
            \cref{case:min-inst-completeness-down}
            to conclude that $[[Γ, nas ⊢ ↓uN SC minby nf(uσ)]]$.
            Then by \ruleref{\ottdruleSINGExistsLabel},
            we have $[[Γ ⊢ ∃nas.↓uN minby nf(uσ)]]$.
    \end{caseof}
\end{proof}

\obsMinInstDeterministic*
\begin{proof}
    We prove it by induction on $[[Γ ⊢ uP SC minby uσ]]$.
    It is easy to see that each inference rule is deterministic.
\end{proof}
